---
title: "Analysis"
---
# Methods 

To compare soundtrack structure across eras, I use: 

- total number of tracks per game 
- total runtime per game 
- average track length 
- track-length distributions 

These measures help show how soundtrack structure expanded as gaming technology evolved

```{r, message=FALSE, warning=FALSE}
library(ggplot2)
library(dplyr)

tracks <- read.csv("video_game_ost_cleaned_with_seconds.csv", stringsAsFactors = FALSE)

game_summary <- tracks |>
  group_by(game, year, era) |>
  summarise(
    total_tracks = n(),
    total_runtime_sec = sum(Duration_sec, na.rm = TRUE),
    avg_track_length = mean(Duration_sec, na.rm = TRUE),
    .groups = "drop"
  ) |>
  mutate(total_runtime_min = total_runtime_sec / 60)

game_summary
```
This table summarizes key structural characteristics of each game's soundtrack. For every game in the dataset, I calculated the total number of tracks, the total runtime of the soundtrack in seconds, and the average length of individual tracks. Converting total runtime into minutes allows easier comparison across titles and eras. These measures provide a baseline view of how soundtrack scale varies across games released at different technological stages of video game history.


## Number of Tracks per Game Over Time

```{r, message=FALSE, warning=FALSE}
ggplot(game_summary, aes(x = year, y = total_tracks)) +
  geom_point(size = 3) +
  geom_smooth(method = "lm", se = FALSE) +
  theme_minimal() +
  labs(
    title = "Number of Tracks per Game Over Time",
    x = "Year",
    y = "Number of Tracks"
  )
```
This plot shows how the number of tracks included in a game's soundtrack changes over time. Each point represents a single game in the dataset, and the fitted trend line illustrates the overall direction of change. Early arcade and 8-bit titles typically contained relatively few tracks because hardware limitations restricted storage and sound channels. As gaming technology advanced, especially with CD-based systems and modern digital storage, soundtracks expanded dramatically. Modern games often include dozens of tracks, reflecting the shift toward more cinematic and atmospheric musical design.

## Runtime by Era

```{r, message=FALSE, warning=FALSE}
ggplot(game_summary, aes(x = era, y = total_runtime_min)) +
  geom_boxplot() +
  theme_minimal() +
  labs(
    title = "Soundtrack Runtime by Era",
    x = "Era",
    y = "Runtime (minutes)"
  )
```

This boxplot compares the total runtime of game soundtracks across technological eras. The distribution highlights how early games tended to have shorter overall soundtracks, often consisting of brief looping musical cues. Later eras show significantly greater runtime variability and longer overall soundtracks. This change reflects the growing importance of music in shaping narrative atmosphere, exploration, and emotional tone in modern video games.

## Distribution of Track Lengths

```{r, message=FALSE, warning=FALSE}
ggplot(tracks, aes(x = Duration_sec)) +
  geom_histogram(bins = 40) +
  theme_minimal() +
  labs(
    title = "Distribution of Track Lengths",
    x = "Track Duration (seconds)",
    y = "Count"
  )
```
The histogram shows the distribution of individual track lengths across the dataset. Track durations vary widely, ranging from very short pieces to extremely long compositions. Many tracks cluster around moderate durations, while a smaller number of outliers represent much longer pieces. This variation suggests that video game soundtracks can include a mix of short looping cues as well as longer compositions, reflecting the diverse musical structures used in game scoring.


## Average Track Length Over Time
```{r}
avg_length_by_year <- game_summary |>
  group_by(year) |>
  summarise(avg_length = mean(avg_track_length, na.rm = TRUE))

ggplot(avg_length_by_year, aes(x = year, y = avg_length)) +
  geom_line(color = "#6aff6a", linewidth = 1) +
  geom_point(color = "#6aff6a", size = 3) +
  geom_smooth(method = "lm", se = FALSE, color = "white") +
  theme_minimal() +
  labs(
    title = "Average Track Length Over Time",
    x = "Year",
    y = "Average Track Length (seconds)"
  )
```
This visualization tracks how the average duration of individual soundtrack pieces changes over time. Early video game music often relied on short looping segments designed to repeat during gameplay, while later games increasingly featured longer compositions. The upward trend suggests a shift away from purely functional loops and toward more expansive soundtrack writing.


