---
title: "Analysis"
---
# Methods 

To compare soundtrack structure across eras, I use: 

- total number of tracks per game 
- total runtime per game 
- average track length 
- track-length distributions 

These measures help show how soundtrack structure expanded as gaming technology evolved

## Nmerical Summary of Data
```{r, message=FALSE, warning=FALSE}
library(ggplot2)
library(dplyr)

df <- read.csv("video_game_ost_cleaned_with_seconds.csv")

game_summary <- df |>
  group_by(game, year, era) |>
  summarise(
    total_tracks = n(),
    total_runtime_sec = sum(Duration_sec, na.rm = TRUE),
    avg_track_length = mean(Duration_sec, na.rm = TRUE),
    .groups = "drop"
  ) |>
  mutate(total_runtime_min = total_runtime_sec / 60)

game_summary
```

## Total Soundtrack Runtime Over Time
```{r, message=FALSE, waring = FALSE}
ggplot(game_summary, aes(x = year, y = total_tracks)) +
  geom_point(size = 3) +
  geom_smooth(method = "lm", se = FALSE) +
  theme_minimal() +
  labs(
    title = "Number of Tracks per Game Over Time",
    x = "Year",
    y = "Number of Tracks"
  )
```


## Runtime by Era
```{r, message=FALSE, warning = FALSE}
ggplot(game_summary, aes(x = era, y = total_runtime_min)) +
  geom_boxplot() +
  theme_minimal() +
  labs(
    title = "Soundtrack Runtime by Era",
    x = "Era",
    y = "Runtime (minutes)"
  )
```


## Distribution of Track Lengths
```{r,message=FALSE, warning=FALSE}
ggplot(df, aes(x = Duration_sec)) +
  geom_histogram(bins = 40) +
  theme_minimal() +
  labs(
    title = "Distribution of Track Lengths",
    x = "Track Duration (seconds)",
    y = "Count"
  )
```

